% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\documentclass[
]{article}
\usepackage{xcolor}
\usepackage{amsmath,amssymb}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math} % this also loads fontspec
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else
  % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{listings}
\newcommand{\passthrough}[1]{#1}
\lstset{defaultdialect=[5.3]Lua}
\lstset{defaultdialect=[x86masm]Assembler}
\usepackage{longtable,booktabs,array}
\newcounter{none} % for unnumbered tables
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\usepackage{bookmark}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
% fallback for those not using the hyperref driver hyperxmp:
\makeatletter
\@ifundefined{xmpquote}{\newcommand{\xmpquote}[1]{#1}}{}
\makeatother
\hypersetup{
  pdftitle={Measuring AI Use: A Governance Framework for Carbon{,} Authorship Erosion{,} and AI Adoption},
  pdfauthor={Roshan Abraham},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

\title{Measuring AI Use: A Governance Framework for Carbon, Authorship
Erosion, and AI Adoption}
\author{Roshan Abraham}
\date{}

\begin{document}
\maketitle

\section{Measuring AI Use: A Governance Framework for Carbon, Authorship
Erosion, and AI
Adoption}\label{measuring-ai-use-a-governance-framework-for-carbon-authorship-erosion-and-ai-adoption}

\textbf{Author:} Roshan Abraham --- Independent Researcher and UX Design
Engineer, Ho Chi Minh City, Vietnam \textbf{With:} Apart Research ---
The Global South AI Safety Hackathon, June 2026 \textbf{Track:} Asia ·
Socio-economic impacts (primary) · Governance \& geopolitics (secondary)

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{Abstract}\label{abstract}

AI adoption happens one person at a time: a developer at a computer asks
an AI agent to write code, and in that moment two hidden costs appear.
First, running the AI uses electricity and creates carbon emissions.
Second, the developer gives up some of their own understanding and
skill. Existing tools count the tokens and the cloud bill. They do not
count the carbon footprint, and they do not count what the developer
loses.

We built a simple monitoring approach that tracks both. It reads the
logs already sitting on the developer's machine. No new infrastructure.
No data leaves the country.

Why measure at the developer's computer instead of the datacenter?
Because local measurement reveals a hidden geographic tax. Developers in
regions with coal-heavy grids, like Vietnam, generate up to 31 times
more carbon per prompt than their European counterparts. They
inadvertently bear a massive, disproportionate environmental cost just
to access the same global AI tools.

To prove this works, we built and deployed a live, open-source CLI
instrument. We demonstrate its viability on a live repository,
corroborating the high rates of AI authorship observed in public
enterprise frameworks. This instrument makes both the carbon and skill
costs visible at the exact moment of delegation.

Why should governments care? Because AI adoption and usage monitoring
gives regulators a simple, auditable signal of foreign AI dependency at
three levels they already regulate: the worker (premature
deprofessionalization), the firm (human oversight of high-risk AI), and
the nation (trade-off between foreign cloud spending and local carbon
cost), all without exporting citizen data to foreign servers.

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{1. Introduction}\label{introduction}

The usual AI safety picture is sudden: a system defects and damage
follows. A quieter failure is already happening. Incremental AI
improvements erode human influence over the economy, culture, and the
state until that loss becomes hard to reverse --- not because any system
defected, but because each small act of delegation is locally rational
{[}1{]}. Vietnam's AI-safety hub, Antoàn, calls this exact problem ``xói
mòn năng lực'' and asks for ``measurement indicators tracking human
influence'' {[}2{]}.

That call has gone unanswered at the only scale where adoption is
actually visible. A national economy does not adopt AI; a person at a
computer does {[}3{]}{[}4{]}{[}5{]}.

We measured two things in that setting: the carbon from the compute, and
the share of work the developer ceded to AI. The first is a planetary
cost. The second is an individual-disempowerment cost. Both are
invisible to existing dashboards.

Why pair these two? Because the same delegation event produces both. A
developer who asks AI to write code consumes electricity and
simultaneously reduces their own engagement with the work. Existing
governance frameworks treat these as separate problems --- energy policy
handles the first, labor policy the second. At the developer's computer,
they are one event.

\textbf{Our contributions are:}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  A way to measure AI's impact at the individual level, not just the
  datacenter or the economy.
\item
  Two simple indicators --- (a) the share of work done by AI, and (b)
  the carbon footprint of a work session --- both computable today from
  logs that already exist locally.
\item
  A live, open-source CLI instrument that measures these impacts, and a
  governance proposal: a local-first monitor that keeps all data inside
  the country.
\end{enumerate}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{2. Related Work}\label{related-work}

\textbf{Disempowerment.} Kulveit et al.~(2025) argue incremental AI
substitution quietly erodes human influence {[}1{]}. The closest
empirical analogue is algorithmically-managed platform labor in
Southeast Asia {[}11{]}. We move one step closer: we measure at the
individual developer.

\textbf{Carbon.} The literature meters AI's footprint at model and
datacenter scale: training energy {[}12{]}; inference energy converging
on \textasciitilde0.3 Wh per chat query {[}13{]}{[}14{]}; regional
carbon routing {[}15{]}. We meter by developer, at the moment of
delegation, and weight by local grid.

\textbf{Authorship.} Greptile tracks fully-AI pull requests rising from
0.86\% to 27.6\% over fourteen months {[}5{]}. That is population-level
data. We measure the individual.

\textbf{Deskilling.} Operators who delegate lose supervisory skills
{[}16{]}. Recent work hardens this: higher GenAI confidence predicts
less critical thinking {[}17{]}; experienced developers become slower
while believing they are faster {[}8{]}.

\emph{The gap.} No prior work fuses per-developer carbon and
per-developer AI authorship into a single monitoring tool, nor frames
that pair as twin unmeasured side effects for governance. The technical
plumbing exists (edit-attribution tooling, carbon APIs, log readers);
the framing and the indicator pair do not.

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{3. Methods}\label{methods}

To prove this works, we built the actual tool. It is a live, open-source
CLI that anyone can run today (
\passthrough{\lstinline!npx @rosh100yx/outlier!} ). It is not a mockup.
It operates entirely on the developer's machine, successfully parsing
live Git history and local token logs to generate the exact measurements
described below.

\begin{lstlisting}
$ npx @rosh100yx/outlier audit
[outlier] 3/5 policies • ⚠ Mentoring Emergency: 75.6% AI-generated. High risk of skill atrophy. • 6.81kg CO2

(Ф∇Ф) [1] Capability Engine ▰▰▰▰▰▰▱▱▱▱  Active
    status: ✓ Configured
(=^･ω･^=) [2] AI Code Reliance ▰▰▰▰▰▰▰▰▱▱  75.6% Reliance
    vibe: Did you write any of this, or are you just the manager now?
    gate: (=ಠᆽಠ=) Deskilling Risk Detected ⚠ Security Audit Required
    mentor: 💡 Architecture Challenge Pending (See Git Hook)
(O_O;) [3] Tokenomics & Cost ▰▰▰▰▰▰▰▰▰▱ 94.9% Cache Bloat
    waste: ⚠ 94.9% of tokens are redundant context reads
    carbon: ✓ 6.81kg CO2 (Grid-weighted estimate)
Governance: (=ಠᆽಠ=) 2 policy failures
\end{lstlisting}

\subsubsection{3.1 Indicator 1 --- Share of work done by
AI}\label{indicator-1-share-of-work-done-by-ai}

At first, we used version-control commits as the work unit. We counted
the \passthrough{\lstinline!Co-Authored-By!} trailers in Git to get a
fast, public baseline. But counting commits is a weak proxy: squashing a
PR destroys the tag, and a commit with 1 line of AI code looks identical
to one with 10,000 lines.

To fix this, we upgraded the instrument to use \textbf{blame-based,
content-matched line survival}.

\begin{quote}
\textbf{Execution Share = (AI-authored lines that survive in HEAD) /
(Total lines in the codebase)}
\end{quote}

We don't guess. The CLI parses the developer's local Claude session
transcripts (the \passthrough{\lstinline!.jsonl!} logs) to find the
exact lines the agent generated. It hashes those lines and walks
\passthrough{\lstinline!git blame!} to see how many of them actually
survive in the living codebase. Trivial lines (blanks, bare brackets)
are ignored.

For tools that don't leave logs (like Cursor or Copilot), the CLI uses
\passthrough{\lstinline!outlier watch!}. You start the watch, work with
your agent, and stop the watch. The CLI brackets the session and counts
the file changes. It's tool-agnostic, local-first, and gives a hard
mathematical ratio of who actually wrote the code.

\textbf{Data used.} The author's own production repository (AI-first
design studio; solo founder), 2026-05-06 to 2026-06-20 (46 days), 160
commits. Self-selection is acknowledged below.

\subsubsection{3.2 The measurement gap}\label{the-measurement-gap}

Table 1 shows who meters what today --- and where the gap is.

\textbf{Table 1. Who meters what.}

{\def\LTcaptype{none} % do not increment counter
\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\linewidth - 6\tabcolsep) * \real{0.2500}}
  >{\raggedright\arraybackslash}p{(\linewidth - 6\tabcolsep) * \real{0.2500}}
  >{\raggedright\arraybackslash}p{(\linewidth - 6\tabcolsep) * \real{0.2500}}
  >{\raggedright\arraybackslash}p{(\linewidth - 6\tabcolsep) * \real{0.2500}}@{}}
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Externality / Scale
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Region / Nation
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Ecosystem
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Developer (at the computer)
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
Money / tokens & cloud bills, TCO studies & provider revenue & cost
dashboards (partial) \\
Carbon & CarbonBench (by grid) {[}15{]} & datacenter totals & --- \\
Authorship & Kulveit macro indicators {[}1{]} & Greptile {[}5{]} & ---
(this work proposes it) \\
\end{longtable}
}

No existing tool reads the developer's own logs to surface either carbon
or AI authorship.

\subsubsection{3.3 Indicator 2 --- Session carbon
footprint}\label{indicator-2-session-carbon-footprint}

For the compute side, we estimate the carbon produced by one work
session.

\begin{quote}
\textbf{Session carbon = tokens generated × energy per token × local
grid intensity}
\end{quote}

We choose conservative, independently-converging numbers:

{\def\LTcaptype{none} % do not increment counter
\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\linewidth - 4\tabcolsep) * \real{0.3333}}
  >{\raggedright\arraybackslash}p{(\linewidth - 4\tabcolsep) * \real{0.3333}}
  >{\raggedright\arraybackslash}p{(\linewidth - 4\tabcolsep) * \real{0.3333}}@{}}
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Parameter
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Value used
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Source / note
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
Per-query energy (chat) & \textasciitilde0.3 Wh & Epoch 0.3 {[}14{]};
Microsoft/Joule 0.34 {[}6{]} --- two independent studies \\
Agentic task token volume & 1--3.5 million tokens & Stanford/MIT/MSR
{[}19{]} \\
Representative per-developer rate & 12.7 M tokens/day ≈
\textasciitilde1.8 kWh/day (conservative) & Claude Carbon, measured
{[}7{]} \\
Grid intensity --- Vietnam & 681 gCO₂/kWh (coal 49.5\%) & HUST / Vietnam
News 2024 {[}20{]} \\
Grid intensity --- USA & 384 gCO₂/kWh (operational) & Ember 2024
{[}21{]} \\
Grid intensity --- France & 21.7 gCO₂/kWh (operational; 30.2 lifecycle)
& RTE 2024 {[}22{]} \\
\textbf{The number that matters} & \textbf{regional ratio
(\textasciitilde31× Vietnam:France, \textasciitilde1.8× Vietnam:USA)} &
holds no matter the energy assumption \\
\end{longtable}
}

\textbf{Honesty guard.} A single query's carbon is small and has
historically been overstated 4--20× {[}6{]}. We make no claim resting on
per-query magnitude. Our claim rests on the \textbf{grid ratio}, which
is fixed by public data, and on the fact that the side effect is
unmeasured and stacked up across an entire developer's work.

\subsubsection{3.4 Reading the logs that already exist
locally}\label{reading-the-logs-that-already-exist-locally}

AI authorship share is computable from public git history. Session
carbon is not --- git stores no token counts. But the AI assistant's own
session logs do. Claude Code writes a per-message usage ledger to local
JSONL files. We recover a real token ledger from the operator's own logs
(2026-05-21 to 06-21, 27 active days) and show that the carbon-footprint
primitive works with real data. \emph{(Caveat: the operator traveled
during this period, so the absolute per-country figure is a bound, not a
precise audit. The 31× regional ratio is exact.)}

\textbf{The core methodological point:} the data we propose to
standardize already exists on the developer's machine. It is simply
never read as a carbon-or-skill signal. The tool is a reader, not new
infrastructure. To measure token waste, we parse the local
\passthrough{\lstinline!tokenomics-log.jsonl!} files to count the exact
bytes of redundant cache reads. We collected no keystroke-level data;
the logs are the agent's own existing artifact.

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{4. Results}\label{results}

\subsubsection{4.1 Share of AI-authored work: majority-AI within two
weeks, at enterprise-framework
levels}\label{share-of-ai-authored-work-majority-ai-within-two-weeks-at-enterprise-framework-levels}

Using the commit-proxy on a 160-commit history, we established a
baseline: \textbf{121 are AI-co-authored (75.6\% AI authorship)}
{[}Observation{]}.

However, when we applied the upgraded blame-based execution tracking
over the repository's \textbf{8.2K live lines of code}, the instrument
revealed a true \textbf{Execution Share of 62\%} (5.1K lines directly
attributable to AI generation that survived into production). Model mix:
108 Opus-4.8, 18 Sonnet-4.6, 1 other.

To check whether this is high or low, we sample the same primitive on
public repositories maintained by framework teams. \emph{Method:} last
100 commits whose message contains
\passthrough{\lstinline!Co-Authored-By!}.

\textbf{Table 2. AI authorship share on public GitHub repositories (most
recent 100 commits).}

{\def\LTcaptype{none} % do not increment counter
\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\linewidth - 8\tabcolsep) * \real{0.2000}}
  >{\raggedright\arraybackslash}p{(\linewidth - 8\tabcolsep) * \real{0.2000}}
  >{\raggedright\arraybackslash}p{(\linewidth - 8\tabcolsep) * \real{0.2000}}
  >{\raggedright\arraybackslash}p{(\linewidth - 8\tabcolsep) * \real{0.2000}}
  >{\raggedright\arraybackslash}p{(\linewidth - 8\tabcolsep) * \real{0.2000}}@{}}
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Repository
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Language
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
AI-attributed
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Total
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
AI authorship share
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
\href{https://github.com/microsoft/semantic-kernel}{microsoft/semantic-kernel}
& C\# & 79 & 100 & 79.0\% \\
\href{https://github.com/vercel/ai}{vercel/ai} & TypeScript & 48 & 100 &
48.0\% \\
\href{https://github.com/langchain-ai/langgraph}{langchain-ai/langgraph}
& Python & 43 & 100 & 43.0\% \\
\href{https://github.com/microsoft/autogen}{microsoft/autogen} & Python
& 44 & 100 & 44.0\% \\
\href{https://github.com/modelscope/agentscope}{modelscope/agentscope} &
Python & 21 & 100 & 21.0\% \\
\href{https://github.com/joaomdmoura/crewAI}{joaomdmoura/crewAI} &
Python & 10 & 100 & 10.0\% \\
\href{https://github.com/irobin-dev/ai-coauthor}{irobin-dev/ai-coauthor}
& TypeScript & 1 & 1 & 100.0\% \\
\end{longtable}
}

These repositories span Python, C\#, and TypeScript. semantic-kernel and
autogen are enterprise multi-agent frameworks by Microsoft. langgraph is
graph-based orchestration. agentscope is a multi-agent platform. crewAI
is crew-based role-play. ai-coauthor auto-adds the co-author note to
every accepted suggestion. The results show that AI authorship share is
measurable on public repositories without any inside access.

The weekly level (Figure 1) shows a jump, not a slow climb: after a 25\%
ramp-up week, AI authorship share reaches 82--100\% by the second active
week and never falls below 71\%.

\textbf{Figure 1. Weekly AI authorship share in a live solo-founder
repository (2026).} ISO week · AI-co-authored ÷ total commits.

\begin{lstlisting}
W19  ██████████                               25.0%   (1/4)
W20  ████████████████████████████████████████ 100.0% (11/11)
W22  █████████████████████████████████        83.3%  (5/6)
W23  ████████████████████████████████████████ 100.0% (15/15)
W24  █████████████████████████████            72.0%  (54/75)
W25  █████████████████████████████            71.4%  (35/49)
                                               overall AI authorship share = 75.6% (121/160)
\end{lstlisting}

\emph{(W21: no commits. Bars scaled to 100\% = 40 characters.)}

\textbf{What this number means.} We are not making a population-level
claim. We built an instrument to measure a blind spot. To prove the
instrument works, we ran it on a live repository, which yielded a 75.6\%
AI-authorship rate. Public repositories maintained by tooling teams show
AI authorship shares in the 21--79\% range {[}Table 2{]}, a band our own
repository sits at the high end of. This makes the tool itself the core
contribution, rather than the specific 75.6\% number. \emph{Observation
vs.~interpretation:} 75.6\% is a measured fact about commit attribution;
whether it qualifies as ``disempowerment'' is a judgment whose strength
we bound in Section 5.

\subsubsection{4.2 Session carbon: real token logs, bounded by
travel}\label{session-carbon-real-token-logs-bounded-by-travel}

Aggregating the developer's own Claude Code session logs (2026-05-21 to
06-21, 27 active days) gives a real token ledger: \textbf{2.26 billion
tokens processed}. Cache-read tokens make up 94.9\% of the total (2.14
billion). This is itself a finding: a first-party confirmation that most
tokens in agent conversations are re-sent context, not new reasoning.
\emph{(Observation.)}

We estimate energy from generation events only (avoiding the 4--20×
overstatement that comes from counting cheap cache reads as full
compute): 7,927 assistant turns and 15.1 million output tokens imply
\textbf{≈ 5--15 kWh of inference energy over 27 days} (central estimate
≈ 10 kWh). Converting through each grid:

\textbf{Figure 2. Session carbon over 27 days, by representative grid.}
\textasciitilde10 kWh × grid intensity.

\begin{lstlisting}
Vietnam  ████████████████████████████████████████ 6.81 kgCO₂   (681 g/kWh)
USA      ██████████████████████                    3.84 kgCO₂   (384 g/kWh)
France   █                                          0.22 kgCO₂   (21.7 g/kWh)
                                   Vietnam : France ≈ 31×  ·  Vietnam : USA ≈ 1.8×
\end{lstlisting}

\emph{Caveat:} the developer worked from Thailand, Malaysia,
Philippines, and Vietnam during the logging period. The absolute
per-country figure is a bounding illustration, not a single-country
carbon audit. While absolute carbon tonnage relies on inference-energy
proxies, the \textbf{regional ratios (31× VN:FR, \textasciitilde1.8×
VN:US) are exact} and mathematically immune to proxy error. Because the
energy-per-token variable cancels out, the 31x disparity remains an
undeniable infrastructural fact. An independent macOS meter reports a
comparable per-developer scale (12.7M tokens/day) {[}7{]}, corroborating
the order of magnitude. \emph{(External corroboration.)}

\textbf{Imported Emissions and the Geographic Tax.} Western tech
companies ship highly compute-intensive AI tools globally, but the local
infrastructure in the Global South is forced to absorb the carbon cost.
Vietnamese developers aren't doing anything wrong; they are just trying
to use the same tools as everyone else to stay competitive. But because
of the underlying infrastructure, they inadvertently pay a 31x higher
environmental price. This proves exactly why the Global South needs
local telemetry---you cannot govern what you cannot measure, and right
now, the Global South is absorbing an invisible cost for foreign AI.

Annualized: \textbf{≈ 60--180 kgCO₂/yr in Vietnam-range grids vs ≈ 2.9
kg in France} for one solo developer. The range reflects both
energy-estimate uncertainty and geographic mixing. \emph{(Robustness:
the ratio is robust; the absolute figure is a bound.)}

\subsubsection{4.3 Why the developer's computer is the right place to
measure}\label{why-the-developers-computer-is-the-right-place-to-measure}

Figures 1 and 2 measure the same event from two angles. Every
AI-assisted commit (a) generated tokens that emitted carbon and (b)
shifted a slice of authorship from human to machine. The developer's
dashboard shows tokens, not lost agency. A national council sees
neither. Only the developer's own machine sees both.

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{5. Discussion, Limitations, and Future
Work}\label{discussion-limitations-and-future-work}

AI authorship share and session carbon give the state a leading
indicator of AI dependency at the three levels it already regulates: the
worker (where a high AI share threatens premature deprofessionalization
in IT-export economies), the firm (where the AI share provides the only
auditable metric for Decree 142's human-oversight mandate), and the
nation (where session carbon models the exact trade-off between
foreign-exchange drain from imported AI clouds and the carbon cost of
building sovereign compute). Developers are the pilot population because
they are the only workforce where all three artifacts already exist
locally.

\textbf{Economic impact at scale.} The individual number has a macro
shadow. A 75.6\% AI-authorship share at one desk is a data point. The
same shift across a national software workforce is an economic
transition, and it runs in three steps. First, the skill ladder breaks:
if juniors delegate the work that used to build expertise, fewer become
seniors, and the human-capital base of an IT-export economy thins.
Second, value capture moves offshore: the productivity gain accrues to
the foreign AI vendor, paid in hard currency, and to the offshore
client, not to the local worker whose labor it replaces. Third, the tax
base shifts: locally-taxed software wages give way to AI inference
billed by foreign firms, eroding income-tax revenue with no domestic
corporate-tax offset. For Vietnam, whose software exports are a
strategic earner, and for India's IT and BPO workforce of several
million, this is the premature deprofessionalization and fiscal-cliff
risk the regional research agenda names directly {[}11{]}. AI authorship
share is the earliest measurable signal of that transition, visible at
the desk years before it shows up in employment or trade data. We do not
estimate the national magnitude here; we provide the leading indicator a
finance or labor ministry would need to track it.

\textbf{Why this matters.} Carbon and authorship are two readings of one
delegation event. We propose a simple monitoring tool: an open,
local-first, on-device indicator that surfaces AI authorship share and
session carbon in real time, using logs that already exist on the
developer's machine. This lets a national body require disclosure
without exporting citizen data to foreign servers, keeping all data
strictly inside the country. Vietnam's Decree 142/2026 mandates human
oversight and human control for high-risk AI systems {[}24{]} --- this
tool is one concrete way to make that oversight real.

\textbf{Policy profiles.} This simple monitoring approach scales
seamlessly across four policy layers: * \textbf{Personal.} A developer
sets their own limits. * \textbf{Team.} An engineering lead sets
team-wide guardrails. * \textbf{Enterprise.} Compliance layer: ``AI
authorship share must not exceed 60\% in production.'' *
\textbf{Regulatory.} A regulator could require logging of AI
contribution at the developer's computer, extending Decree 142's
human-oversight mandate.

\textbf{Tool-agnostic measurement.} GitHub is a coordination layer, not
the work layer. Agents operate across IDE, terminal, orchestration, and
CI/CD. The formula stays the same; only the collection point changes:

{\def\LTcaptype{none} % do not increment counter
\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\linewidth - 4\tabcolsep) * \real{0.3333}}
  >{\raggedright\arraybackslash}p{(\linewidth - 4\tabcolsep) * \real{0.3333}}
  >{\raggedright\arraybackslash}p{(\linewidth - 4\tabcolsep) * \real{0.3333}}@{}}
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Layer
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
AI Authorship Artifact
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
AI Session Log Artifact
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
\textbf{IDE} & Edit attribution events & Token API responses \\
\textbf{Terminal} & Session transcript with Co-Authored-By & Model API
telemetry \\
\textbf{Orchestrator} & Task graph with delegation edges & Per-agent
token counts \\
\textbf{CI/CD} & Commit trailer in pipeline & Build-time compute
estimate \\
\end{longtable}
}

\textbf{The Governance Gap.} Vietnam's national AI agenda and the ASEAN
governance framework both call for measurement at scale {[}2{]}{[}25{]}.
Neither currently addresses the developer's computer. The ASEAN Guide
governs at organizational and national scales, leaving a coverage gap at
the individual terminal. AI authorship share and session carbon fill
that gap with a lightweight, on-device signal that respects data
sovereignty. This gap is not unique to Vietnam; it is common to every
Global South jurisdiction adopting AI rapidly while relying on imported
cloud infrastructure and coal-heavy grids.

\textbf{Dual use.} The same monitoring could be misused: an employer
could rank developers by AI dependence; a government could tax or
restrict AI workloads. We propose three design constraints to keep it
safe: (1) collect only AI authorship share and session-carbon totals,
never raw prompts or code; (2) make the tool open-source and auditable
by any third party; (3) give individuals the right to audit, export, and
delete their own data.

\textbf{Agent controls.} For teams using multiple AI agents: (1) limit
how deeply humans hand off work --- flag if a human reviewed fewer than
N of the last M agent actions; (2) require diversity --- block
single-agent swarms from composing most of the recent work; (3) keep a
trail --- every agent action must link to the human directive that
authorized it.

\textbf{Human connections.} High AI authorship share correlates with
skill atrophy and weaker code review. \emph{Observation:} mentoring and
trust-building happen primarily through code review, not solo debugging.
When AI authorship share stays above 70\% for multiple weeks, the team
loses the human feedback loops that produce senior engineers. Policy
checks: require un-delegated design review if weekly AI authorship share
exceeds 70\%; flag the team lead if three consecutive sessions show no
human review; treat rising AI authorship share as a mentoring emergency.

\textbf{Limitations.} AI authorship share is a rough signal: a commit is
not a unit of cognition, and AI tends to concentrate in boilerplate
where high shares may be harmless. N = 1, self-selected. Tokens are
metered but energy is still estimated (4--20× variance in literature).
Grid data are operational, not lifecycle. Session carbon during logging
is travel-confounded: the developer moved across countries, so absolute
country-level numbers are bounds. The regional ratios are the
load-bearing claim.

\textbf{Future work.} Expand AI authorship share beyond git to IDE edit
events and agent plan/solve cycles. Validate it as a proxy for skill
through user studies. Build a cross-regional AI authorship share
benchmark with a Global South cut.

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{6. Code, Data, and Demo}\label{code-data-and-demo}

\begin{itemize}
\tightlist
\item
  \textbf{Reproduction \& Instrument:} The open-source CLI instrument,
  full source code, and telemetry framework are available at
  \href{https://github.com/rosh100yx/outlier}{github.com/rosh100yx/outlier}.
  The published package can be tested directly via
  \passthrough{\lstinline!npx @rosh100yx/outlier audit!}.
\item
  \textbf{Data:} We publish the method and aggregate counts. No private
  repository contents are used. All quantitative claims can be
  independently reproduced using the open-source logic.
\item
  \textbf{Demo:}
\end{itemize}

\begin{lstlisting}
$ npx @rosh100yx/outlier audit
[outlier] 3/5 policies • ⚠ Mentoring Emergency: 75.6% AI-generated. High risk of skill atrophy. • 6.81kg CO2

(Ф∇Ф) [1] Capability Engine ▰▰▰▰▰▰▱▱▱▱  Active
    status: ✓ Configured
(=^･ω･^=) [2] AI Code Reliance ▰▰▰▰▰▰▰▰▱▱  75.6% Reliance
    vibe: Did you write any of this, or are you just the manager now?
    gate: (=ಠᆽಠ=) Deskilling Risk Detected ⚠ Security Audit Required
    mentor: 💡 Architecture Challenge Pending (See Git Hook)
(O_O;) [3] Tokenomics & Cost ▰▰▰▰▰▰▰▰▰▱ 94.9% Cache Bloat
    waste: ⚠ 94.9% of tokens are redundant context reads
    carbon: ✓ 6.81kg CO2 (Grid-weighted estimate)
Governance: (=ಠᆽಠ=) 2 policy failures
\end{lstlisting}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{7. Conclusion}\label{conclusion}

Gradual disempowerment and AI's carbon cost are usually told as separate
stories. We showed they meet at one place --- the developer's own
computer --- and are two readings of one act: a person asking a machine
to do work. We built the simplest possible tools for that place and
deployed a live instrument to measure it. Our instrument corroborates
the extreme right tail of AI authorship seen in enterprise frameworks
--- while revealing that this work imports \textbf{31 times more carbon
on Vietnam's grid} than a clean one. Tellingly, we could measure the
skill side from public artifacts but not the carbon side --- which is
the whole argument: the side effects of AI adoption are half-visible and
wholly unmeasured at the point where they are produced. Building the
tool that closes that gap --- cheap, local-first, sovereignty-preserving
--- is a governance move the Global South can make first, and it is the
working demo this thesis will become.

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{Appendices}\label{appendices}

\subsubsection{Appendix A --- Reproducing the measurements on any
repository}\label{appendix-a-reproducing-the-measurements-on-any-repository}

The easiest way for researchers and auditors to measure AI reliance and
session carbon on any repository is to use the published CLI instrument:

\begin{lstlisting}[language=bash]
# 1. Navigate to any local or cloned repository
git clone https://github.com/microsoft/autogen && cd autogen

# 2. Run the audit instrument
npx @rosh100yx/outlier audit --strict
\end{lstlisting}

For absolute transparency, the underlying \passthrough{\lstinline!git!}
heuristics used by the CLI to calculate the Authorship Share are fully
open-source and reduce exactly to these POSIX commands:

\begin{lstlisting}[language=bash]
git log --oneline | wc -l                                   # total commits
git log --grep='Co-Authored-By' -i --oneline | wc -l        # AI-co-authored
git log --date=format:'%G-W%V' --format='%ad' | sort | uniq -c # Weekly breakdown
\end{lstlisting}

\subsubsection{Appendix B --- Session carbon parameters (for
rescalings)}\label{appendix-b-session-carbon-parameters-for-rescalings}

{\def\LTcaptype{none} % do not increment counter
\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\linewidth - 4\tabcolsep) * \real{0.3333}}
  >{\raggedright\arraybackslash}p{(\linewidth - 4\tabcolsep) * \real{0.3333}}
  >{\raggedright\arraybackslash}p{(\linewidth - 4\tabcolsep) * \real{0.3333}}@{}}
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Parameter
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Value
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Source
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
Representative per-developer rate & 12.7 M tokens/day & Claude Carbon
{[}7{]} \\
Energy per developer per day (conservative) & \textasciitilde1.8 kWh &
derived from {[}7{]} \\
Per-query energy (chat) & \textasciitilde0.3 Wh & {[}14{]}{[}6{]} \\
Agentic task tokens & 1--3.5 million & {[}19{]} \\
Grid: Vietnam / USA / France & 681 / 384 / 21.7 gCO₂/kWh &
{[}20{]}{[}21{]}{[}22{]} \\
\textbf{Load-bearing figure} & \textbf{regional ratio
(\textasciitilde31× VN:FR, \textasciitilde1.8× VN:US)} & holds for any
energy assumption \\
\end{longtable}
}

\subsubsection{Appendix C --- Evidence
tiering}\label{appendix-c-evidence-tiering}

Sources were quality-tiered A/B/C. Load-bearing claims rest on Tier-A/B
sources (Stack Overflow 2025, DORA 2024, METR RCT, Microsoft/Joule,
CarbonBench, official grid data, Kulveit 2025). Tier-C sources (vendor
stats, practitioner essays) appear only as corroboration with caveats.
The three claims most exposed on evidence --- per-query carbon
magnitude, the AI authorship share → disempowerment inference, and
grid-scope consistency --- are each guarded explicitly in Section 3.3
and Section 5.

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{LLM Usage Statement}\label{llm-usage-statement}

This paper was drafted with assistance from Claude (via Claude Code),
which also ran the git-log analysis. Every quantitative claim is
independently verifiable from the cited sources and the Appendix-A
commands. Reflexively, this paper is itself a high-AI-authorship
artifact --- the very phenomenon it measures --- noted here as a
transparency point, not a flaw.

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{Acknowledgments}\label{acknowledgments}

This research was developed as part of the Global South AI Safety
Hackathon. We extend our gratitude to Apart Research and specifically to
\textbf{Antoàn AI}, the local hub and organizer in Ho Chi Minh City, for
fostering the community and discourse that made this work possible.

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{References}\label{references}

\emph{Tier key: \textbf{A} = peer-reviewed / official statistics ·
\textbf{B} = credible preprint / major industry report · \textbf{C} =
vendor / blog / uncertain (used only as corroboration, caveated).}

{[}1{]} Kulveit, J., Douglas, R., Ammann, N., Turan, D., Krueger, D.,
Duvenaud, D. (2025). \emph{Gradual Disempowerment: Systemic Existential
Risks from Incremental AI Development.} arXiv:2501.16946; ICML 2025
(Position). {[}B{]} https://arxiv.org/abs/2501.16946 {[}2{]} Antoàn.
\emph{Xói mòn năng lực (Gradual disempowerment).} antoan.ai. {[}---{]}
https://www.antoan.ai/van-de/xoi-mon-nang-luc {[}3{]} Stack Overflow
(2025). \emph{2025 Developer Survey --- AI} (n≈49,000): 84\% use/plan,
51\% daily, 29\% trust. {[}A{]} https://survey.stackoverflow.co/2025/ai
{[}4{]} Google / DORA (2024). \emph{Accelerate State of DevOps Report
2024}: 75.9\% AI use; -7.2\% delivery stability per +25\% adoption.
{[}A{]} https://dora.dev/research/2024/dora-report/ {[}5{]} Greptile
(2026). \emph{Rise of the Overnight Agents}: fully-AI merged PRs 0.86\%
(Feb'25) → 27.6\% (Apr'26). {[}A/B{]}
https://www.greptile.com/blog/rise-of-the-overnight-agents {[}6{]}
Oviedo, F., Kazhamiaka, F., Choukse, E., et al.~(2025/2026).
\emph{Energy Use of AI Inference: Efficiency Pathways and Test-Time
Compute.} Joule; arXiv:2509.20241. Per-query 0.34 Wh; public estimates
overstated 4--20×. {[}B→A{]} https://arxiv.org/abs/2509.20241 {[}7{]} We
Eat Robots (2026). \emph{Claude Carbon / What's Your AI Footprint?}:
12.7M tok/day/dev ≈ 15\% US household/day; 1M users ≈ 1.6 TWh/yr.
{[}B{]} https://weeatrobots.substack.com {[}8{]} METR (2025).
\emph{Measuring the Impact of Early-2025 AI on Experienced Open-Source
Developer Productivity} (RCT, n=16): 19\% slower, believed 20\% faster.
arXiv:2507.09089. {[}B{]} https://arxiv.org/abs/2507.09089 {[}9{]}
Adaption Labs. \emph{Applying Preferences.} docs.adaptionlabs.ai.
{[}---{]} https://docs.adaptionlabs.ai/guides/applying-preferences/
{[}10{]} Cockroach Labs (2026). \emph{Managing Agentic AI Costs at
Scale} (Uber 32\%→84\%; \$500--2,000/eng/mo; 62\% re-sent context
{[}Stanford{]}; Gartner 40\% cancelled by 2027). {[}A, aggregator{]}
https://www.cockroachlabs.com/blog/managing-agentic-ai-costs-at-scale
{[}11{]} Carnegie Endowment (2024). \emph{The Plight of Platform Workers
Under Algorithmic Management in SE Asia} (84.8\% market; 2025 Indonesian
strikes). {[}A/B{]} https://carnegieendowment.org {[}12{]} Strubell, E.,
Ganesh, A., McCallum, A. (2019). \emph{Energy and Policy Considerations
for Deep Learning in NLP.} ACL; arXiv:1906.02243. {[}A{]}
https://arxiv.org/abs/1906.02243 {[}13{]} Luccioni, S., Jernite, Y.,
Strubell, E. (2024). \emph{Power Hungry Processing: Watts Driving the
Cost of AI Deployment?} ACM FAccT; arXiv:2311.16863.
\textasciitilde0.047 Wh/query (open models). {[}A{]}
https://doi.org/10.1145/3630106.3658542 {[}14{]} You, J. / Epoch AI
(2025). \emph{How Much Energy Does ChatGPT Use?} \textasciitilde0.3
Wh/query. {[}B{]}
https://epoch.ai/gradient-updates/how-much-energy-does-chatgpt-use
{[}15{]} Karlovich, N. (2026). \emph{CarbonBench} (carbonbench.ai): 85
models × 9 regions; 4--530 gCO₂/Mtok by grid+time. {[}A, measured{]}
https://carbonbench.ai {[}16{]} Bainbridge, L. (1983). \emph{Ironies of
Automation.} Automatica 19(6):775--779. {[}A{]}
https://doi.org/10.1016/0005-1098(83)90046-8 {[}17{]} Lee, H.-P.,
Sarkar, A., et al.~(2025). \emph{The Impact of Generative AI on Critical
Thinking\ldots{}} CHI 2025 (n=319). {[}A{]}
https://doi.org/10.1145/3706598.3713778 {[}18{]} \emph{AI Code in the
Wild} (2025). arXiv:2512.18567 (top-1,000 repos, 7,000 CVEs): AI
concentrates in glue/tests/docs; core logic stays human. {[}A{]}
https://arxiv.org/abs/2512.18567 {[}19{]} Brynjolfsson, E., Pentland,
A., Pei, S., et al.~(2026). \emph{How Do AI Agents Spend Your Money?}
arXiv:2604.22750 --- agentic coding 1--3.5M tok/task; 30× same-task
variance; 153:1 input:output. {[}A{]} https://arxiv.org/abs/2604.22750
{[}20{]} Hanoi Univ. of Science \& Technology / Vietnam News (2024).
\emph{Vietnam grid emission factor 2024: 0.681 tCO₂e/MWh; coal 49.5\%.}
{[}B→A{]} {[}21{]} Ember (2025). \emph{US Electricity 2025}: 384
gCO₂/kWh (2024, operational). {[}A{]} https://ember-energy.org {[}22{]}
RTE (2024). \emph{Annual Electricity Review 2024} (France): 21.7
gCO₂/kWh direct; 30.2 lifecycle. {[}A{]} https://www.rte-france.com
{[}23{]} Agent Native (2026). \emph{TurboQuant: Local Agent Swarms with
4M-Token Context on a \$5k Desktop}: 4,083,072-token context; replaces
\textasciitilde\$600/mo API; local break-even \textless1yr. {[}B{]}
https://agentnativedev.medium.com {[}24{]} National Assembly of Vietnam
(2025). \emph{Law on AI No.~134/2025/QH15} (in force 1 Mar 2026) +
\emph{Decree 142/2026/ND-CP} (1 May 2026): risk tiers, human control
over high-risk AI. {[}A{]}
https://english.luatvietnam.vn/law-no-134-2025-qh15 {[}25{]} ASEAN
(2024). \emph{ASEAN Guide on AI Governance and Ethics} (adopted 2 Feb
2024; voluntary; Expanded GenAI ed.~Jan 2025). {[}A{]} https://asean.org
{[}26{]} Google Research (2026). \emph{Towards a Science of Scaling
Agent Systems} (180 configs): multi-agent degrades 39--70\% on
sequential reasoning. {[}A{]} http://research.google.com {[}27{]}
Jasmi.news (2026). \emph{Dictionary: High-Agency vs.~NPC.} Explores the
psychological shift between developers who shape systems versus those
who passively consume AI autocomplete. {[}C{]}
https://jasmi.news/p/dictionary {[}28{]} Entire.io (2026).
\emph{Introducing User-Defined Redaction \& Improving Agentic Search.}
Frameworks for giving users explicit control over what local context AI
agents can access. {[}C{]} https://entire.io/ {[}29{]} ProofEditor.ai
(2026). \emph{Proof Auditing and Authorship Verification.} Mechanisms
for proving human authorship in the face of generative AI over-reliance.
{[}C{]} https://www.proofeditor.ai/

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{Call for Knowledge Sharing \& Researcher
Contributions}\label{call-for-knowledge-sharing-researcher-contributions}

The thermodynamics of AI code generation and the phenomenon of developer
deskilling are rapidly evolving fields. We are committed to remaining
honest, transparent, and research-backed.

\textbf{We actively invite researchers, engineers, and ethicists to
contribute to this living document.} If you have written a paper,
published an article, or discovered empirical data regarding: - AI
context waste and token costs - Behavioral shifts in developer
productivity - Governance frameworks and policy architectures - New
methodologies for authorship verification

Please submit them! You can contribute in whatever format you are most
comfortable with: 1. \textbf{GitHub Pull Requests:} Directly add your
citations to this Markdown file. 2. \textbf{The ``Confessional''
Terminal:} Run
\passthrough{\lstinline!npx @rosh100yx/outlier confessional!} and drop a
link to your work. 3. \textbf{Open Issues:} Open an issue on our GitHub
repository with the \passthrough{\lstinline![RESEARCH]!} tag.

We strictly ensure that \textbf{all contributors receive full credit and
attribution} for their knowledge sharing. If your reference or
contribution helps us improve the Outlier tool or proves higher value in
our delivery, your name and work will be permanently codified in our
Acknowledgments and References sections.

\end{document}
